---
title: "STAD70_Strategy_2_Final"
output:
  pdf_document: default
  html_document: default
date: "2026-04-05"
---

``````{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
Sys.setlocale("LC_ALL", "English")
```

```{r}
rm(list = ls())

library(tidyverse)
library(lubridate)
library(zoo)
library(here)

Sys.setenv(TZ = "America/New_York")

spy_df <- read_csv(here("spy_etf_train.csv")) %>%
  mutate(timestamp  = with_tz(timestamp, tzone = "America/New_York")) %>%
  arrange(timestamp) %>%
  mutate(trade_date = as.Date(timestamp, tz = "America/New_York"),
         log_ret    = log(close / open))

opt_df <- read_csv(here("spy_opt_train.csv")) %>%
  mutate(timestamp  = with_tz(timestamp, tzone = "America/New_York")) %>%
  arrange(timestamp) %>%
  mutate(trade_date = as.Date(timestamp, tz = "America/New_York"))

all_days <- sort(unique(spy_df$trade_date))
cat("Total trading days:", length(all_days), "\n")
```

```{r}
TRAIN_DAYS <- 252
train_days <- all_days[1:TRAIN_DAYS]
test_days  <- all_days[(TRAIN_DAYS + 1):length(all_days)]

cat("Training  :", length(train_days), "days |",
    as.character(min(train_days)), "to", as.character(max(train_days)), "\n")
cat("Test (OOS):", length(test_days),  "days |",
    as.character(min(test_days)),  "to", as.character(max(test_days)),  "\n")
```

```{r}
get_opt_price <- function(type_char, strike_val, ts, tol_mins = 5) {
  ts_end <- ts + minutes(tol_mins)
  tmp <- opt_df %>%
    filter(type == type_char, strike == strike_val,
           timestamp >= ts, timestamp <= ts_end) %>%
    arrange(timestamp)
  if (nrow(tmp) == 0) return(NA_real_)
  px <- tmp$vwap[1]              # VWAP first  <-- FIX 1
  if (is.na(px)) px <- tmp$open[1]
  as.numeric(px)
}

apply_slippage <- function(price, legs = 1, side = c("sell", "buy"),
                           slippage_per_leg = 0.02) {
  side <- match.arg(side)
  adj  <- slippage_per_leg * legs
  if (side == "sell") return(max(price - adj, 0))
  return(price + adj)
}
```

```{r}
compute_daily_features <- function(days, check_time, vol_lookback) {

  feats <- map_dfr(days, function(d) {
    day_data <- spy_df %>% filter(trade_date == d) %>% arrange(timestamp)
    if (nrow(day_data) < 200) {
      return(tibble(trade_date = d,
                    sigma15 = NA_real_, sigma60 = NA_real_,
                    vol15   = NA_real_, vol60   = NA_real_,
                    spy_ref = NA_real_, ts_ref  = as.POSIXct(NA)))
    }

    sigma15 <- rollapply(day_data$log_ret, width = 15, FUN = sd,  fill = NA, align = "right")
    sigma60 <- rollapply(day_data$log_ret, width = 60, FUN = sd,  fill = NA, align = "right")
    vol15   <- rollapply(day_data$volume,  width = 15, FUN = sum, fill = NA, align = "right")
    vol60   <- rollapply(day_data$volume,  width = 60, FUN = sum, fill = NA, align = "right")

    day_data <- day_data %>%
      mutate(sigma15 = sigma15, sigma60 = sigma60, vol15 = vol15, vol60 = vol60)

    target_ts <- ymd_hms(paste(d, check_time), tz = "America/New_York")
    idx <- which.min(abs(as.numeric(day_data$timestamp) - as.numeric(target_ts)))

    tibble(trade_date = d,
           sigma15 = day_data$sigma15[idx],
           sigma60 = day_data$sigma60[idx],
           vol15   = day_data$vol15[idx],
           vol60   = day_data$vol60[idx],
           spy_ref = day_data$open[idx],
           ts_ref  = day_data$timestamp[idx])
  })

  feats %>%
    arrange(trade_date) %>%
    mutate(med30_sigma60 = rollapply(sigma60, width = vol_lookback,
                                     FUN = function(x) median(x, na.rm = TRUE),
                                     fill = NA, align = "right"))
}
```

```{r}
base_feats <- compute_daily_features(all_days,
                                     check_time   = "11:00:00",
                                     vol_lookback = 30)

iv_proxy <- map_dfr(all_days, function(d) {
  entry_ts <- ymd_hms(paste(d, "11:01:00"), tz = "America/New_York")
  feats    <- base_feats %>% filter(trade_date == d)
  if (nrow(feats) == 0 || is.na(feats$spy_ref)) return(NULL)

  strikes_today <- opt_df %>% filter(trade_date == d) %>% distinct(strike) %>% pull(strike)
  if (length(strikes_today) == 0) return(NULL)
  K <- strikes_today[which.min(abs(strikes_today - feats$spy_ref))]

  call_px <- get_opt_price("C", K, entry_ts, tol_mins = 5)
  put_px  <- get_opt_price("P", K, entry_ts, tol_mins = 5)
  if (any(is.na(call_px), is.na(put_px))) return(NULL)

  tibble(trade_date  = d,
         straddle_px = call_px + put_px,
         spy_ref     = feats$spy_ref,
         iv_proxy    = (call_px + put_px) / feats$spy_ref)
}) %>%
  arrange(trade_date) %>%
  mutate(med30_iv_proxy = rollapply(iv_proxy, width = 30,
                                    FUN = function(x) median(x, na.rm = TRUE),
                                    fill = NA, align = "right"),
         iv_rich = iv_proxy > 1.10 * med30_iv_proxy)

cat("Days where IV proxy > 1.10 x 30-day median (options 'rich'):\n")
print(iv_proxy %>% filter(!is.na(iv_rich)) %>% count(iv_rich))
```

```{r}
simulate_day <- function(d, daily_feats,
                         CHECK_TIME, ENTRY_TIME, TIME_EXIT, FINAL_TIME,
                         VOL_COMPRESS_MULT, VOL_EXPAND_MULT,
                         STOP_MULT, SLIPPAGE_PER_LEG) {

  feats <- daily_feats %>% filter(trade_date == d)
  if (nrow(feats) == 0) return(NULL)
  if (any(is.na(feats$sigma15), is.na(feats$sigma60), is.na(feats$med30_sigma60),
          is.na(feats$vol15),   is.na(feats$vol60),   is.na(feats$spy_ref))) return(NULL)

  condA <- feats$sigma15 < VOL_COMPRESS_MULT * feats$sigma60
  condB <- feats$sigma60 < feats$med30_sigma60
  if (!(condA && condB)) return(NULL)

  entry_ts     <- ymd_hms(paste(d, ENTRY_TIME), tz = "America/New_York")
  time_exit_ts <- ymd_hms(paste(d, TIME_EXIT),  tz = "America/New_York")
  final_ts     <- ymd_hms(paste(d, FINAL_TIME), tz = "America/New_York")

  strikes_today <- opt_df %>% filter(trade_date == d) %>% distinct(strike) %>% pull(strike)
  if (length(strikes_today) == 0) return(NULL)
  K <- strikes_today[which.min(abs(strikes_today - feats$spy_ref))]

  call_entry_raw <- get_opt_price("C", K, entry_ts, tol_mins = 5)
  put_entry_raw  <- get_opt_price("P", K, entry_ts, tol_mins = 5)
  if (any(is.na(call_entry_raw), is.na(put_entry_raw))) return(NULL)

  call_entry      <- apply_slippage(call_entry_raw, legs = 1, side = "sell",
                                    slippage_per_leg = SLIPPAGE_PER_LEG)
  put_entry       <- apply_slippage(put_entry_raw,  legs = 1, side = "sell",
                                    slippage_per_leg = SLIPPAGE_PER_LEG)
  premium         <- call_entry + put_entry   
  premium_dollars <- 100 * premium             

  day_data <- spy_df %>%
    filter(trade_date == d, timestamp >= entry_ts) %>%
    arrange(timestamp) %>%
    mutate(sigma15 = rollapply(log_ret, width = 15, FUN = sd, fill = NA, align = "right"),
           sigma60 = rollapply(log_ret, width = 60, FUN = sd, fill = NA, align = "right"))
  if (nrow(day_data) < 50) return(NULL)

  for (i in seq_len(nrow(day_data))) {
    ts <- day_data$timestamp[i]
    S  <- day_data$close[i]

    if (ts >= time_exit_ts) {
      call_exit_raw <- get_opt_price("C", K, ts, tol_mins = 5)
      put_exit_raw  <- get_opt_price("P", K, ts, tol_mins = 5)
      if (!any(is.na(call_exit_raw), is.na(put_exit_raw))) {
        call_exit <- apply_slippage(call_exit_raw, legs = 1, side = "buy",
                                    slippage_per_leg = SLIPPAGE_PER_LEG)
        put_exit  <- apply_slippage(put_exit_raw,  legs = 1, side = "buy",
                                    slippage_per_leg = SLIPPAGE_PER_LEG)
        exit_cost <- 100 * (call_exit + put_exit)
        return(tibble(trade_date = d, strike = K,
                      entry_ts = entry_ts, exit_ts = ts,
                      entry_premium = premium, exit_cost = exit_cost / 100,
                      pnl = premium_dollars - exit_cost,
                      exit_reason = "TIME_EXIT"))
      }
    }

    if (!is.na(S) && abs(S - K) > STOP_MULT * premium) {
      call_exit_raw <- get_opt_price("C", K, ts, tol_mins = 5)
      put_exit_raw  <- get_opt_price("P", K, ts, tol_mins = 5)
      if (!any(is.na(call_exit_raw), is.na(put_exit_raw))) {
        call_exit <- apply_slippage(call_exit_raw, legs = 1, side = "buy",
                                    slippage_per_leg = SLIPPAGE_PER_LEG)
        put_exit  <- apply_slippage(put_exit_raw,  legs = 1, side = "buy",
                                    slippage_per_leg = SLIPPAGE_PER_LEG)
        exit_cost <- 100 * (call_exit + put_exit)
        return(tibble(trade_date = d, strike = K,
                      entry_ts = entry_ts, exit_ts = ts,
                      entry_premium = premium, exit_cost = exit_cost / 100,
                      pnl = premium_dollars - exit_cost,
                      exit_reason = "STOP_LOSS"))
      }
    }

    s15 <- day_data$sigma15[i]
    s60 <- day_data$sigma60[i]
    if (!is.na(s15) && !is.na(s60) && s15 > VOL_EXPAND_MULT * s60) {
      call_exit_raw <- get_opt_price("C", K, ts, tol_mins = 5)
      put_exit_raw  <- get_opt_price("P", K, ts, tol_mins = 5)
      if (!any(is.na(call_exit_raw), is.na(put_exit_raw))) {
        call_exit <- apply_slippage(call_exit_raw, legs = 1, side = "buy",
                                    slippage_per_leg = SLIPPAGE_PER_LEG)
        put_exit  <- apply_slippage(put_exit_raw,  legs = 1, side = "buy",
                                    slippage_per_leg = SLIPPAGE_PER_LEG)
        exit_cost <- 100 * (call_exit + put_exit)
        return(tibble(trade_date = d, strike = K,
                      entry_ts = entry_ts, exit_ts = ts,
                      entry_premium = premium, exit_cost = exit_cost / 100,
                      pnl = premium_dollars - exit_cost,
                      exit_reason = "VOL_EXPANSION"))
      }
    }

    if (ts >= final_ts) break
  }

  close_row <- spy_df %>%
    filter(trade_date == d) %>%
    arrange(timestamp) %>%
    filter(format(timestamp, "%H:%M:%S") == FINAL_TIME)
  if (nrow(close_row) == 0)
    close_row <- spy_df %>% filter(trade_date == d) %>%
      arrange(timestamp) %>% slice_tail(n = 1)

  S_T         <- close_row$close[1]
  call_payoff <- max(S_T - K, 0)
  put_payoff  <- max(K - S_T, 0)
  loss        <- call_payoff + put_payoff

  tibble(trade_date = d, strike = K,
         entry_ts = entry_ts, exit_ts = final_ts,
         entry_premium = premium, exit_cost = loss,
         pnl = 100 * (premium - loss),
         exit_reason = "HOLD_TO_EXPIRY")
}

run_backtest <- function(days, check_time, entry_time, time_exit,
                         vol_compress_mult, vol_expand_mult,
                         stop_mult, vol_lookback, slippage_per_leg) {
  feats   <- compute_daily_features(days, check_time, vol_lookback)
  results <- map_dfr(days, ~simulate_day(
    .x, feats,
    CHECK_TIME        = check_time,
    ENTRY_TIME        = entry_time,
    TIME_EXIT         = time_exit,
    FINAL_TIME        = "15:59:00",
    VOL_COMPRESS_MULT = vol_compress_mult,
    VOL_EXPAND_MULT   = vol_expand_mult,
    STOP_MULT         = stop_mult,
    SLIPPAGE_PER_LEG  = slippage_per_leg
  ))
  if (is.null(results) || nrow(results) == 0) return(NULL)
  results %>% arrange(trade_date) %>% mutate(cum_pnl = cumsum(pnl))
}
```


```{r cache=TRUE}
# param_grid <- expand.grid(
#   CHECK_TIME        = c("10:30:00", "11:00:00", "11:30:00"),  
#   VOL_COMPRESS_MULT = c(0.70, 0.80, 0.90),                    
#   VOL_LOOKBACK_DAYS = c(20L, 30L),                             
#   STOP_MULT         = c(0.80, 0.90),                           
#   TIME_EXIT         = c("14:15:00", "14:30:00", "14:45:00"), 
#   VOL_EXPAND_MULT   = c(1.50, 2.00),                           
#   stringsAsFactors  = FALSE
# )%>%
#   mutate(ENTRY_TIME = format(
#     as.POSIXct(CHECK_TIME, format = "%H:%M:%S") + 60, "%H:%M:%S"
#   ))
# 
# cat("Total combinations:", nrow(param_grid), "\n")
# 
# grid_results <- map_dfr(seq_len(nrow(param_grid)), function(i) {
#   p   <- param_grid[i, ]
#   res <- run_backtest(
#     days              = train_days,
#     check_time        = p$CHECK_TIME,
#     entry_time        = p$ENTRY_TIME,
#     time_exit         = p$TIME_EXIT,
#     vol_compress_mult = p$VOL_COMPRESS_MULT,
#     vol_expand_mult   = p$VOL_EXPAND_MULT,
#     stop_mult         = p$STOP_MULT,
#     vol_lookback      = p$VOL_LOOKBACK_DAYS,
#     slippage_per_leg  = 0.02
#   )
#   if (is.null(res) || nrow(res) == 0) {
#     return(tibble(i = i, n_trades = 0L, total_pnl = NA_real_,
#                   sharpe = NA_real_, win_rate = NA_real_, max_dd = NA_real_))
#   }
#   mu <- mean(res$pnl); s <- sd(res$pnl)
#   dd <- min(res$cum_pnl - cummax(res$cum_pnl))
#   tibble(i         = i,
#          n_trades  = nrow(res),
#          total_pnl = sum(res$pnl),
#          sharpe    = if (!is.na(s) && s > 0) mu / s * sqrt(252) else NA_real_,
#          win_rate  = mean(res$pnl > 0),
#          max_dd    = dd)
# })
# 
# grid_summary <- bind_cols(param_grid, grid_results %>% select(-i)) %>%
#   arrange(desc(total_pnl))
# 
# cat("\n=== Top 10 Configurations by Training P&L ===\n")
# grid_summary %>%
#   select(CHECK_TIME, VOL_COMPRESS_MULT, VOL_LOOKBACK_DAYS,
#          STOP_MULT, TIME_EXIT, VOL_EXPAND_MULT,
#          n_trades, total_pnl, sharpe, win_rate, max_dd) %>%
#   head(10) %>% print()
```

```{r}
# best <- grid_summary %>%
#   filter(!is.na(total_pnl), n_trades >= 5) %>%
#   slice(1)
# 
# CHECK_TIME        <- best$CHECK_TIME
# ENTRY_TIME        <- best$ENTRY_TIME
# TIME_EXIT         <- best$TIME_EXIT
# FINAL_TIME        <- "15:59:00"          
# VOL_COMPRESS_MULT <- best$VOL_COMPRESS_MULT
# VOL_EXPAND_MULT   <- best$VOL_EXPAND_MULT
# STOP_MULT         <- best$STOP_MULT
# VOL_LOOKBACK_DAYS <- best$VOL_LOOKBACK_DAYS
# SLIPPAGE_PER_LEG  <- 0.02              
# 
# cat("=== Best Hyperparameters (selected on training P&L) ===\n")
# cat("CHECK_TIME        :", CHECK_TIME,        "\n")
# cat("ENTRY_TIME        :", ENTRY_TIME,        "\n")
# cat("VOL_COMPRESS_MULT :", VOL_COMPRESS_MULT, "\n")
# cat("VOL_LOOKBACK_DAYS :", VOL_LOOKBACK_DAYS, "\n")
# cat("STOP_MULT         :", STOP_MULT,         "\n")
# cat("TIME_EXIT         :", TIME_EXIT,         "\n")
# cat("VOL_EXPAND_MULT   :", VOL_EXPAND_MULT,   "\n")
# cat("SLIPPAGE_PER_LEG  :", SLIPPAGE_PER_LEG,  "\n")
```
```{r}
# === Best Hyperparameters (selected on training P&L) ===
# CHECK_TIME        : 11:00:00 
# ENTRY_TIME        : 11:01:00 
# VOL_COMPRESS_MULT : 0.9 
# VOL_LOOKBACK_DAYS : 20 
# STOP_MULT         : 0.8 
# TIME_EXIT         : 14:30:00 
# VOL_EXPAND_MULT   : 1.5 
# SLIPPAGE_PER_LEG  : 0.02 

CHECK_TIME = "11:00:00"
ENTRY_TIME = "11:01:00"
TIME_EXIT = "14:30:00"
VOL_COMPRESS_MULT = 0.9
VOL_EXPAND_MULT = 1.5
STOP_MULT = 0.8
VOL_LOOKBACK_DAYS = 20
SLIPPAGE_PER_LEG = 0.02


FINAL_TIME        <- "15:59:00"          

```

```{r}
train_results <- run_backtest(
  days              = train_days,
  check_time        = CHECK_TIME,
  entry_time        = ENTRY_TIME,
  time_exit         = TIME_EXIT,
  vol_compress_mult = VOL_COMPRESS_MULT,
  vol_expand_mult   = VOL_EXPAND_MULT,
  stop_mult         = STOP_MULT,
  vol_lookback      = VOL_LOOKBACK_DAYS,
  slippage_per_leg  = SLIPPAGE_PER_LEG
)
```

```{r}
trade_table <- train_results %>%
  transmute(
    Date                = trade_date,
    Strike              = strike,
    Entry_Time          = format(entry_ts, "%H:%M"),
    Exit_Time           = format(exit_ts,  "%H:%M"),
    Entry_Premium       = round(entry_premium, 4),
    Exit_Cost_or_Payoff = round(exit_cost,     4),
    PnL                 = round(pnl,     2),
    Cum_PnL             = round(cum_pnl, 2),
    Exit_Reason         = exit_reason
  )

print(trade_table)
```

```{r}
ggplot(train_results, aes(x = trade_date, y = cum_pnl)) +
  geom_line(color = "steelblue", linewidth = 1) +
  geom_point(aes(color = pnl > 0), size = 2) +
  scale_color_manual(values = c("TRUE"  = "darkgreen", "FALSE" = "darkred"),
                     labels  = c("TRUE"  = "Win",       "FALSE" = "Loss"),
                     name    = "") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  labs(
    title    = "In-Sample Cumulative P&L: Short ATM 0DTE Straddle",
    subtitle = paste0("Best params from grid search | Training period: ",
                      min(train_days), " to ", max(train_days)),
    x        = "Date",
    y        = "Cumulative P&L ($)",
    caption  = paste0("Trades: ", nrow(train_results),
                      "  |  Total P&L: $", round(sum(train_results$pnl), 0),
                      "  |  Win Rate: ",
                      round(mean(train_results$pnl > 0) * 100, 1), "%")
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "top", plot.caption = element_text(size = 8))
```

```{r}
final_stats <- train_results %>%
  summarise(
    n_trades     = n(),
    total_pnl    = round(sum(pnl), 2),
    avg_pnl      = round(mean(pnl), 2),
    sd_pnl       = round(sd(pnl), 2),
    win_rate     = round(mean(pnl > 0), 3),
    payoff_ratio = round(mean(pnl[pnl > 0]) / abs(mean(pnl[pnl < 0])), 3),
    med_pnl      = round(median(pnl), 2),
    p05          = round(quantile(pnl, 0.05), 2),
    p95          = round(quantile(pnl, 0.95), 2),
    max_drawdown = round(min(cum_pnl - cummax(cum_pnl)), 2),
    sharpe       = round(mean(pnl) / sd(pnl) * sqrt(252), 3)
  )

print(final_stats)
print(train_results %>% count(exit_reason))
```

```{r}
test_results <- run_backtest(
  days              = test_days,
  check_time        = CHECK_TIME,
  entry_time        = ENTRY_TIME,
  time_exit         = TIME_EXIT,
  vol_compress_mult = VOL_COMPRESS_MULT,
  vol_expand_mult   = VOL_EXPAND_MULT,
  stop_mult         = STOP_MULT,
  vol_lookback      = VOL_LOOKBACK_DAYS,
  slippage_per_leg  = SLIPPAGE_PER_LEG
)
```

```{r}
ggplot(test_results, aes(x = trade_date, y = cum_pnl)) +
  geom_line(color = "tomato", linewidth = 1) +
  geom_point(aes(color = pnl > 0), size = 2) +
  scale_color_manual(values = c("TRUE"  = "darkgreen", "FALSE" = "darkred"),
                     labels  = c("TRUE"  = "Win",       "FALSE" = "Loss"),
                     name    = "") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  labs(
    title    = "Out-of-Sample Cumulative P&L: Short ATM 0DTE Straddle",
    subtitle = paste0("Fixed params from training | Test period: ",
                      min(test_days), " to ", max(test_days)),
    x        = "Date",
    y        = "Cumulative P&L ($)",
    caption  = paste0("Trades: ", nrow(test_results),
                      "  |  Total P&L: $", round(sum(test_results$pnl), 0),
                      "  |  Win Rate: ",
                      round(mean(test_results$pnl > 0) * 100, 1), "%")
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "top", plot.caption = element_text(size = 8))
```

```{r}
oos_stats <- test_results %>%
  summarise(
    n_trades     = n(),
    total_pnl    = round(sum(pnl), 2),
    avg_pnl      = round(mean(pnl), 2),
    sd_pnl       = round(sd(pnl), 2),
    win_rate     = round(mean(pnl > 0), 3),
    payoff_ratio = round(mean(pnl[pnl > 0]) / abs(mean(pnl[pnl < 0])), 3),
    med_pnl      = round(median(pnl), 2),
    p05          = round(quantile(pnl, 0.05), 2),
    p95          = round(quantile(pnl, 0.95), 2),
    max_drawdown = round(min(cum_pnl - cummax(cum_pnl)), 2),
    sharpe       = round(mean(pnl) / sd(pnl) * sqrt(252), 3)
  )

print(oos_stats)
print(test_results %>% count(exit_reason))
```
