---
title: "STAD70_Strategy_2_Final"
output:
  pdf_document: default
  html_document: default
date: "2026-04-05"
---

``````{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)
library(zoo)
library(ggplot2)
library(here)

Sys.setenv(TZ = "America/New_York")

#parameters from grid search on training data

CHECK_TIME        <- "11:00:00"
ENTRY_TIME        <- "11:01:00"
TIME_EXIT         <- "14:30:00"
FINAL_TIME        <- "15:59:00"
VOL_COMPRESS_MULT <- 0.9
VOL_EXPAND_MULT   <- 1.5
STOP_MULT         <- 0.8
VOL_LOOKBACK_DAYS <- 20L
SLIPPAGE_PER_LEG  <- 0.02


# spy_path = "spy_etf_train.csv"
# opt_path = "spy_opt_train.csv"

spy_path = "spy_etf_test.csv"
opt_path = "spy_opt_test.csv"


#load test data
spy_df <- read_csv(here(spy_path)) %>%
  mutate(timestamp  = with_tz(timestamp, tzone = "America/New_York")) %>%
  arrange(timestamp) %>%
  mutate(trade_date = as.Date(timestamp, tz = "America/New_York"),
         log_ret    = log(close / open))

opt_df <- read_csv(here(opt_path)) %>%
  mutate(timestamp  = with_tz(timestamp, tzone = "America/New_York")) %>%
  arrange(timestamp) %>%
  mutate(trade_date = as.Date(timestamp, tz = "America/New_York"))

all_days <- sort(unique(spy_df$trade_date))
cat("Test days:", length(all_days), "\n")

get_opt_price <- function(type_char, strike_val, ts, tol_mins = 5) {
  ts_end <- ts + minutes(tol_mins)
  tmp <- opt_df %>%
    filter(type == type_char, strike == strike_val,
           timestamp >= ts, timestamp <= ts_end) %>%
    arrange(timestamp)
  if (nrow(tmp) == 0) return(NA_real_)
  px <- tmp$vwap[1]
  if (is.na(px)) px <- tmp$open[1]
  as.numeric(px)
}

apply_slippage <- function(price, legs = 1, side = c("sell", "buy")) {
  side <- match.arg(side)
  adj  <- SLIPPAGE_PER_LEG * legs
  if (side == "sell") return(max(price - adj, 0))
  return(price + adj)
}

compute_daily_features <- function(days) {
  feats <- map_dfr(days, function(d) {
    day_data <- spy_df %>% filter(trade_date == d) %>% arrange(timestamp)
    if (nrow(day_data) < 200) {
      return(tibble(trade_date = d,
                    sigma15 = NA_real_, sigma60 = NA_real_,
                    vol15 = NA_real_, vol60 = NA_real_,
                    spy_ref = NA_real_, ts_ref = as.POSIXct(NA)))
    }
    sigma15 <- rollapply(day_data$log_ret, width = 15, FUN = sd,  fill = NA, align = "right")
    sigma60 <- rollapply(day_data$log_ret, width = 60, FUN = sd,  fill = NA, align = "right")
    vol15   <- rollapply(day_data$volume,  width = 15, FUN = sum, fill = NA, align = "right")
    vol60   <- rollapply(day_data$volume,  width = 60, FUN = sum, fill = NA, align = "right")
    day_data <- day_data %>%
      mutate(sigma15 = sigma15, sigma60 = sigma60, vol15 = vol15, vol60 = vol60)
    target_ts <- ymd_hms(paste(d, CHECK_TIME), tz = "America/New_York")
    idx <- which.min(abs(as.numeric(day_data$timestamp) - as.numeric(target_ts)))
    tibble(trade_date = d,
           sigma15 = day_data$sigma15[idx], sigma60 = day_data$sigma60[idx],
           vol15   = day_data$vol15[idx],   vol60   = day_data$vol60[idx],
           spy_ref = day_data$open[idx],    ts_ref  = day_data$timestamp[idx])
  })
  feats %>%
    arrange(trade_date) %>%
    mutate(med30_sigma60 = rollapply(sigma60, width = VOL_LOOKBACK_DAYS,
                                     FUN = function(x) median(x, na.rm = TRUE),
                                     fill = NA, align = "right"))
}

simulate_day <- function(d, daily_feats) {
  feats <- daily_feats %>% filter(trade_date == d)
  if (nrow(feats) == 0) return(NULL)
  if (any(is.na(feats$sigma15), is.na(feats$sigma60), is.na(feats$med30_sigma60),
          is.na(feats$vol15), is.na(feats$vol60), is.na(feats$spy_ref))) return(NULL)

  # entry signal
  condA <- feats$sigma15 < VOL_COMPRESS_MULT * feats$sigma60
  condB <- feats$sigma60 < feats$med30_sigma60
  if (!(condA && condB)) return(NULL)

  entry_ts     <- ymd_hms(paste(d, ENTRY_TIME), tz = "America/New_York")
  time_exit_ts <- ymd_hms(paste(d, TIME_EXIT),  tz = "America/New_York")
  final_ts     <- ymd_hms(paste(d, FINAL_TIME), tz = "America/New_York")

  # pick ATM strike
  strikes_today <- opt_df %>% filter(trade_date == d) %>% distinct(strike) %>% pull(strike)
  if (length(strikes_today) == 0) return(NULL)
  K <- strikes_today[which.min(abs(strikes_today - feats$spy_ref))]

  # entry prices (vwap)
  call_entry_raw <- get_opt_price("C", K, entry_ts)
  put_entry_raw  <- get_opt_price("P", K, entry_ts)
  if (any(is.na(call_entry_raw), is.na(put_entry_raw))) return(NULL)

  call_entry <- apply_slippage(call_entry_raw, legs = 1, side = "sell")
  put_entry  <- apply_slippage(put_entry_raw,  legs = 1, side = "sell")
  premium         <- call_entry + put_entry
  premium_dollars <- 100 * premium

  # intraday loop
  day_data <- spy_df %>%
    filter(trade_date == d, timestamp >= entry_ts) %>%
    arrange(timestamp) %>%
    mutate(sigma15 = rollapply(log_ret, width = 15, FUN = sd, fill = NA, align = "right"),
           sigma60 = rollapply(log_ret, width = 60, FUN = sd, fill = NA, align = "right"))
  if (nrow(day_data) < 50) return(NULL)

  for (i in seq_len(nrow(day_data))) {
    ts <- day_data$timestamp[i]
    S  <- day_data$close[i]

    # time exit
    if (ts >= time_exit_ts) {
      ce_raw <- get_opt_price("C", K, ts)
      pe_raw <- get_opt_price("P", K, ts)
      if (!any(is.na(ce_raw), is.na(pe_raw))) {
        ce <- apply_slippage(ce_raw, 1, "buy")
        pe <- apply_slippage(pe_raw, 1, "buy")
        exit_cost <- 100 * (ce + pe)
        return(tibble(trade_date = d, strike = K,
                      entry_ts = entry_ts, exit_ts = ts,
                      entry_premium = premium, exit_cost = exit_cost / 100,
                      call_entry_vwap = call_entry_raw, put_entry_vwap = put_entry_raw,
                      call_exit_vwap  = ce_raw,         put_exit_vwap  = pe_raw,
                      pnl = premium_dollars - exit_cost, exit_reason = "TIME_EXIT"))
      }
    }

    # stop loss
    if (!is.na(S) && abs(S - K) > STOP_MULT * premium) {
      ce_raw <- get_opt_price("C", K, ts)
      pe_raw <- get_opt_price("P", K, ts)
      if (!any(is.na(ce_raw), is.na(pe_raw))) {
        ce <- apply_slippage(ce_raw, 1, "buy")
        pe <- apply_slippage(pe_raw, 1, "buy")
        exit_cost <- 100 * (ce + pe)
        return(tibble(trade_date = d, strike = K,
                      entry_ts = entry_ts, exit_ts = ts,
                      entry_premium = premium, exit_cost = exit_cost / 100,
                      call_entry_vwap = call_entry_raw, put_entry_vwap = put_entry_raw,
                      call_exit_vwap  = ce_raw,         put_exit_vwap  = pe_raw,
                      pnl = premium_dollars - exit_cost, exit_reason = "STOP_LOSS"))
      }
    }

    # vol expansion exit
    s15 <- day_data$sigma15[i]; s60 <- day_data$sigma60[i]
    if (!is.na(s15) && !is.na(s60) && s15 > VOL_EXPAND_MULT * s60) {
      ce_raw <- get_opt_price("C", K, ts)
      pe_raw <- get_opt_price("P", K, ts)
      if (!any(is.na(ce_raw), is.na(pe_raw))) {
        ce <- apply_slippage(ce_raw, 1, "buy")
        pe <- apply_slippage(pe_raw, 1, "buy")
        exit_cost <- 100 * (ce + pe)
        return(tibble(trade_date = d, strike = K,
                      entry_ts = entry_ts, exit_ts = ts,
                      entry_premium = premium, exit_cost = exit_cost / 100,
                      call_entry_vwap = call_entry_raw, put_entry_vwap = put_entry_raw,
                      call_exit_vwap  = ce_raw,         put_exit_vwap  = pe_raw,
                      pnl = premium_dollars - exit_cost, exit_reason = "VOL_EXPANSION"))
      }
    }

    if (ts >= final_ts) break
  }

  # hold to expiry
  close_row <- spy_df %>%
    filter(trade_date == d) %>%
    arrange(timestamp) %>%
    filter(format(timestamp, "%H:%M:%S") == FINAL_TIME)
  if (nrow(close_row) == 0)
    close_row <- spy_df %>% filter(trade_date == d) %>% arrange(timestamp) %>% slice_tail(n = 1)

  S_T         <- close_row$close[1]
  call_payoff <- max(S_T - K, 0)
  put_payoff  <- max(K - S_T, 0)

  tibble(trade_date = d, strike = K,
         entry_ts = entry_ts, exit_ts = final_ts,
         entry_premium = premium, exit_cost = call_payoff + put_payoff,
         call_entry_vwap = call_entry_raw, put_entry_vwap = put_entry_raw,
         call_exit_vwap  = call_payoff,    put_exit_vwap  = put_payoff,
         pnl = 100 * (premium - call_payoff - put_payoff),
         exit_reason = "HOLD_TO_EXPIRY")
}

daily_feats <- compute_daily_features(all_days)

trade_results <- map_dfr(all_days, ~simulate_day(.x, daily_feats)) %>%
  arrange(trade_date) %>%
  mutate(cum_pnl = cumsum(pnl))

cat("Trades:", nrow(trade_results), "\n")
cat("Total P&L: $", round(sum(trade_results$pnl), 2), "\n")

cashflow_rows <- map_dfr(seq_len(nrow(trade_results)), function(i) {
  r <- trade_results[i, ]
  K <- r$strike

  entry_call <- tibble(
    timestamp  = r$entry_ts,
    asset      = paste0("Call K=", K),
    position   = -1,
    vwap_price = r$call_entry_vwap,
    cashflow   = 100 * apply_slippage(r$call_entry_vwap, 1, "sell")
  )
  entry_put <- tibble(
    timestamp  = r$entry_ts,
    asset      = paste0("Put K=", K),
    position   = -1,
    vwap_price = r$put_entry_vwap,
    cashflow   = 100 * apply_slippage(r$put_entry_vwap, 1, "sell")
  )

  exit_call <- tibble(
    timestamp  = r$exit_ts,
    asset      = paste0("Call K=", K),
    position   = +1,
    vwap_price = r$call_exit_vwap,
    cashflow   = -100 * apply_slippage(r$call_exit_vwap, 1, "buy")
  )
  exit_put <- tibble(
    timestamp  = r$exit_ts,
    asset      = paste0("Put K=", K),
    position   = +1,
    vwap_price = r$put_exit_vwap,
    cashflow   = -100 * apply_slippage(r$put_exit_vwap, 1, "buy")
  )

  bind_rows(entry_call, entry_put, exit_call, exit_put)
})

cashflow_table <- cashflow_rows %>%
  arrange(timestamp) %>%
  mutate(timestamp = format(timestamp, "%Y-%m-%d %H:%M"),
         cum_pnl = cumsum(cashflow))

write_csv(cashflow_table, "cashflow_table_Short_Strad.csv")
cat("Saved cashflow_table.csv\n")

# cumulative P&L plot
p <- ggplot(trade_results, aes(x = trade_date, y = cum_pnl)) +
  geom_line(color = "steelblue", linewidth = 1) +
  geom_point(aes(color = pnl > 0), size = 2) +
  scale_color_manual(values = c("TRUE" = "darkgreen", "FALSE" = "darkred"),
                     labels  = c("TRUE" = "Win", "FALSE" = "Loss"), name = "") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  labs(title   = "Out-of-Sample Cumulative P&L: Short ATM 0DTE Straddle",
       x       = "Date",
       y       = "Cumulative P&L ($)",
       caption = paste0("Trades: ", nrow(trade_results),
                        " | Total P&L: $", round(sum(trade_results$pnl), 0),
                        " | Win Rate: ", round(mean(trade_results$pnl > 0) * 100, 1), "%")) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "top", plot.caption = element_text(size = 8))

ggsave("cum_pnl_plot_Short_Strad.png", plot = p, width = 10, height = 5, dpi = 150)
cat("Saved cum_pnl_plot.png\n")

```